Flow polytopes

### 2019-05-17 pages/math/semi-comb/project/paper.tex
Lazy update

\documentclass{amsart}
\usepackage[T1]{fontenc}
%\usepackage{mathpazo}
%\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
%\usepackage{rotating}
%\usepackage{color}
%\usepackage{tabu}
%\usepackage{parskip}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{calc}
\usepackage{xfrac}
\usepackage{mathtools}
\usepackage{cancel}
\usepackage{framed}
\usepackage{nth}
\usepackage{tikz}
\usetikzlibrary{calc,decorations.markings}
%\usepackage{xcolor-solarized}
%\usepackage{minted}
%\usepackage{MnSymbol}
%\usepackage[libertine]{newtxmath}
%\newcommand*\circled[1]{\tikz[baseline= (char.base)]{
    %\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
%\usepackage{xypic}
%\usepackage[tightpage]{preview} % display
% Math operators
\DeclareMathOperator{\Z}{\mathbb{Z}}
\DeclareMathOperator{\Q}{\mathbb{Q}}
\DeclareMathOperator{\N}{\mathbb{N}}
\DeclareMathOperator{\R}{\mathbb{R}}
\DeclareMathOperator{\C}{\mathbb{C}}
\DeclareMathOperator{\F}{\mathbb{F}}
\DeclareMathOperator{\sge}{\mathsf{e}}
\DeclareMathOperator{\sgm}{\mathsf{m}}
\DeclareMathOperator{\sgG}{\mathsf{G}}
\DeclareMathOperator{\sgg}{\mathsf{g}}
\DeclareMathOperator{\sgF}{\mathsf{F}}
\DeclareMathOperator{\sgt}{\mathsf{t}}
\DeclareMathOperator{\sgL}{\mathsf{L}}
\DeclareMathOperator{\sgZ}{\mathsf{Z}}
\DeclareMathOperator{\sgAp}{Ap}
\DeclareMathOperator{\sgPF}{PF}
\DeclareMathOperator{\sgconv}{conv}
\DeclareMathOperator{\sgsupp}{supp}
\DeclareMathOperator{\calC}{\mathcal{C}}
\DeclareMathOperator{\calF}{\mathcal{F}}
\DeclareMathOperator{\calO}{\mathcal{O}}
\DeclareMathOperator{\Grho}{\prescript{}{\rho}G}
% Theorem environments
\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem*{thm*}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem*{lemma*}{Lemma}
\newtheorem*{claim}{Claim}
\newtheorem*{cor}{Corollary}
\newtheorem*{prop}{Proposition}
\newtheorem*{conj}{Conjecture}
\theoremstyle{remark}
\newtheorem*{example*}{Example}
\theoremstyle{definition}
\newtheorem*{definition*}{Definition}
% Special theorem displays
\newenvironment{example}%
	{\begin{leftbar}\begin{example*}
}{%
	\end{example*}\end{leftbar}
}
\newenvironment{definition}%
	{\begin{leftbar}\begin{definition*}
}{%
	\end{definition*}\end{leftbar}
}
% Paired math operators
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
% Math helper functions
\newcommand{\congr}[2]{{[#1]}_{#2}}
\newcommand{\set}[1]{\{#1\}}
\newcommand{\genset}[1]{{\langle{}#1\rangle{}}}
\newcommand{\abs}[1]{\lvert{}#1\rvert{}}
\newcommand{\norm}[1]{\lVert{}#1\rVert{}}
\newcommand{\bmx}[1]{\begin{bmatrix}#1\end{bmatrix}}
\newcommand{\pmx}[1]{\begin{pmatrix}#1\end{pmatrix}}
  \newcommand{\sbmx}[1]{{\left[\begin{smallmatrix}#1\end{smallmatrix}\right]}}
\newcommand{\smallxymatrix}[2]{\xymatrix@C-=#1in@R-=#1in{#2}}
\renewcommand\labelitemi{\small$\bullet$}
%
%\makeatletter
\tikzset{%
  gdot/.style={draw,shape=circle,fill=black,inner sep=1},
  ddot/.style={draw,shape=circle,fill=white,inner sep=1},
  dedge/.style={draw=black!40},
  midarrow/.style={postaction={decorate},decoration={markings,mark=at position 0.5 with {\arrow{>}}}}
}
%\makeatother
\begin{document}
%
%
%

\section*{Prompt:}

@@

Graphs are a major area of concern in mathematics, having an entire field to
study dedicated to them. Many different families of graphs exist, but for this
paper we'll concern ourselves with directed graphs and $\Z_n$-flows on these
graphs, all concepts which the first section introduces. The second section
is dedicated to a reciprocity expression for counting the number of possible
flows on directed graphs, and to developing some insight into it and its proof.

@@

\section{Introduction to graphs and flows}

@@

A graph $G=(V,E)$ is a collection of $V$, a set of vertices, and $E$, a set of
edges between the vertices.
Now, a graph doesn't have to have any vertices, nor
does it have to to have any edges (except that if it has edges there must be
vertices to which those edges connect).
Also note that the nodes of a graph can be elements of some arbitrary
d-dimensional space, but for this report we'll only concern ourselves with
graphs in some two-dimensional space.
(Note that the word ``node'' is analogous to ``vertex'' and that this paper
will do plenty of haphazard switching between the two of them.)
\begin{definition}[Component, connected component]
  A \emph{connected component} (or simply a \emph{component}) of a graph $G$ is
  a subgraph (a subset of the vertices and edges of $G$) in which any two
  vertices are either directly connected via an edge or are connected to
  one-another via a \emph{path} (a series of vertices and edges). A single
  vertex with no connecting edges is in itself a connected component.
\end{definition}
Any graph then is the disjoint union of connected components.
For example Figure~\ref{fig:disjoint-union} can be called a single graph that is the disjoint
union of three distinct connected components.
\begin{figure}[ht]
  \begin{tikzpicture}[every node/.style=gdot]
    \path
      (.5,.5) node (a) {}
      (0,0) node (b) {} (1,0) node (c) {} (1,1) node (d) {} (0,1) node (e) {}
      (2,1) node (f) {} (2,0) node (g) {} (3,0) node (h) {};
    \draw (a);
    \draw (b) -- (c) -- (d) -- (e) -- (b);
    \draw (f) -- (g) -- (h);
  \end{tikzpicture}
  \caption[Disjoin union]{}
  \label{fig:disjoint-union}
\end{figure}
\begin{lemma*}
  A connected component of the graph $G$ is a maximal subgraph of $G$ in which
  any two nodes are connect by a path.
\end{lemma*}
\begin{lemma*}
  A graph $G$ is \emph{connected} if it has only one connected component.
\end{lemma*}
\begin{definition}[Bridge]
  A \emph{bridge} is a single edge which serves as a connection between two connected
  components which, if removed, would increase the number of connected
  components of the graph by one.
\end{definition}
For example, removing the bridge in Figure~\ref{fig:bridge} turns a single
connected component into two triangular connected components:
one complete graph into a graph with two complete subgraphs.
\begin{figure}[ht]
  \begin{tikzpicture}[every node/.style=gdot]
    \path
      (-.5,.5) node (a) {} ($({-(1+sqrt(3))/2},0)$) node (b) {} ($({-(1+sqrt(3))/2},1)$) node (c) {}
      ( .5,.5) node (d) {} ($({ (1+sqrt(3))/2},0)$) node (e) {} ($({ (1+sqrt(3))/2},1)$) node (f) {};
    \draw (a) -- (b) -- (c) -- (a);
    \draw (d) -- (e) -- (f) -- (d);
    \draw[dashed] (a) -- (d);
  \end{tikzpicture}
  \hspace{1in}
  \begin{tikzpicture}[every node/.style=gdot]
    \path
      (-.5,.5) node (a) {} ($({-(1+sqrt(3))/2},0)$) node (b) {} ($({-(1+sqrt(3))/2},1)$) node (c) {}
      ( .5,.5) node (d) {} ($({ (1+sqrt(3))/2},0)$) node (e) {} ($({ (1+sqrt(3))/2},1)$) node (f) {};
    \draw (a) -- (b) -- (c) -- (a);
    \draw (d) -- (e) -- (f) -- (d);
  \end{tikzpicture}
  \caption[Bridges]{A graph with and without a bridge.}
  \label{fig:bridge}
\end{figure}

@@

So far we've only looked at \emph{unoriented} graphs, but
now we introduce a notion of direction, or orientation, to graphs.
\begin{definition}[Directed, oriented graph]
  A \emph{directed} or \emph{oriented graph} is a graph on which we
  define an \emph{orientation} via a subset $\rho$ of the edges $E$ of the
  graph. For each edge $e=v_i v_j\in E$ (connecting the vertices $v_i,v_j\in
  V$) with $i<j$ we direct:
  \[
    v_i\stackrel{e}{\leftarrow}v_j\text{ if }e\in\rho\text{, and }
    v_i\stackrel{e}{\rightarrow}v_j\text{ if }e\notin\rho.
  \]
\end{definition}
This definition may be confusing at first glance because $\rho$
may be missing some of the edges of $E$. What this definition attempts to succinctly
encapsulate is only the edges with their direction pointing from the vertex of higher
index to the vertex of lower index (which does certainly feel backwards),
with all the remaining edges not in $\rho$ directed in the opposite direction,
from lower to higher index.
Figure~\ref{fig:directed-graph} tries to illustrate this encapsulation for a graph
with five vertices and an orientation $\rho=\{14,23,24\}$.
\begin{figure}[ht]
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style=gdot]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw (v_4) -- (v_3) -- (v_2) -- (v_1) -- (v_4) -- (v_2);
    \end{tikzpicture}
    \caption{The undirected graph.}
  \end{subfigure}
  %\hspace{1in}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style=gdot]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw[midarrow] (v_4) -- (v_1);
      \draw[midarrow] (v_4) -- (v_2);
      \draw[midarrow] (v_3) -- (v_2);
      \draw (v_3) -- (v_4);
      \draw (v_1) -- (v_2);
    \end{tikzpicture}
    \caption{$\rho$ edges marked.}
  \end{subfigure}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style={gdot}]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw[midarrow] (v_4) -- (v_1);
      \draw[midarrow] (v_4) -- (v_2);
      \draw[midarrow] (v_3) -- (v_2);
      \draw[midarrow] (v_3) -- (v_4);
      \draw[midarrow] (v_1) -- (v_2);
    \end{tikzpicture}
    \caption{$E\setminus\rho$ edges marked.}
  \end{subfigure}
  \caption[A directed graph]{A directed graph.}
  \label{fig:directed-graph}
\end{figure}

As an aside, directed graphs are the subject of many algorithms in computer
science, often pertaining to finding optimal ways to traverse the paths of said
graphs. In some models each edge is considered to be of equal ``weight'' as any
other. That is if two edges originating from the same node both point to the
same node, there should be no real preference to traveling down one edge or the
other to get to the destination. On the other hand we might assign a weight to
each edge, presenting a cost-benefit analysis in choosing a path to take (as in
Figure~\ref{fig:path-finding} where the edge with weight 5 now becomes most
expensive to traverse versus the one with weight 2).
\begin{figure}[ht]
  \begin{tikzpicture}
    \path (0,0) node[gdot] (a){} (2,0) node[gdot] (b){};
    \draw[midarrow] (a) to [out=45, in=135] node[above]{} (b);
    \draw[midarrow] (a) to [out=315,in=225] node[below]{$\phantom 5$} (b);
  \end{tikzpicture}
  \hspace{1in}
  \begin{tikzpicture}
    \path (0,0) node[gdot] (a){} (2,0) node[gdot] (b){};
    \draw[midarrow] (a) to [out=45, in=135] node[above]{$2$} (b);
    \draw[midarrow] (a) to [out=315,in=225] node[below]{$5$} (b);
  \end{tikzpicture}
  \caption{A directed graph with and without weights.}
  \label{fig:path-finding}
\end{figure}
Intuitively these weights are often simple $\Z$ integers.

@@

  \caption{Various graphs and their duals.}
  \label{fig:graphs-and-duals}

@@

% ------------------------------------------------------------------------------
% 7.6
% ------------------------------------------------------------------------------
\pagebreak
\section{Reciprocity in counting graph flows}

Ultimately the purpose of this report is in providing some intuition into the
following theorem and its proof:
\begin{thm}\label{thm:1.2.5}
  Let $G$ be a bridgeless graph. For every positive integer $n$, the
  reciprocity statement:
  \[
    {(-1)}^{\xi(G)}\varphi_G(-n)
  \]
  counts the number of pairs $(f,p)$,
  where $f$ is a $\Z_n$-flow and $\rho$ is a totally-cyclic reorientation of
  $G/\sgsupp(f)$. In particular, when $n=-1$, the statement equals the
  number of totally-cyclic orientations of $G$.
\end{thm}

Going forward we'll attempt to build some insight into theorem~\ref{thm:1.2.5}
and its proof, and into the propositions upon which the proof relies.

\subsection{Flows and flow spaces}

>>>>

\documentclass{amsart}
\usepackage[T1]{fontenc}
%\usepackage{mathpazo}
%\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
%\usepackage{rotating}
%\usepackage{color}
%\usepackage{tabu}
%\usepackage{parskip}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{calc}
\usepackage{xfrac}
\usepackage{mathtools}
\usepackage{cancel}
\usepackage{framed}
\usepackage{nth}
\usepackage{tikz}
\usetikzlibrary{calc,decorations.markings}
%\usepackage{xcolor-solarized}
%\usepackage{minted}
%\usepackage{MnSymbol}
%\usepackage[libertine]{newtxmath}
%\newcommand*\circled[1]{\tikz[baseline= (char.base)]{
    %\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
%\usepackage{xypic}
%\usepackage[tightpage]{preview} % display
% Math operators
\DeclareMathOperator{\Z}{\mathbb{Z}}
\DeclareMathOperator{\Q}{\mathbb{Q}}
\DeclareMathOperator{\N}{\mathbb{N}}
\DeclareMathOperator{\R}{\mathbb{R}}
\DeclareMathOperator{\C}{\mathbb{C}}
\DeclareMathOperator{\F}{\mathbb{F}}
\DeclareMathOperator{\sfe}{\mathsf{e}}
\DeclareMathOperator{\sfg}{\mathsf{g}}
\DeclareMathOperator{\sfm}{\mathsf{m}}
\DeclareMathOperator{\sft}{\mathsf{t}}
\DeclareMathOperator{\sfF}{\mathsf{F}}
\DeclareMathOperator{\sfG}{\mathsf{G}}
\DeclareMathOperator{\sfL}{\mathsf{L}}
\DeclareMathOperator{\sfP}{\mathsf{P}}
\DeclareMathOperator{\sfZ}{\mathsf{Z}}
\DeclareMathOperator{\sgAp}{Ap}
\DeclareMathOperator{\sgPF}{PF}
\DeclareMathOperator{\sgehr}{ehr}
\DeclareMathOperator{\sgconv}{conv}
\DeclareMathOperator{\sgsupp}{supp}
\DeclareMathOperator{\calC}{\mathcal{C}}
\DeclareMathOperator{\calF}{\mathcal{F}}
\DeclareMathOperator{\calO}{\mathcal{O}}
\DeclareMathOperator{\Grho}{\prescript{}{\rho}G}
% Theorem environments
\theoremstyle{plain}
\newtheorem{THM}{Theorem}
\newtheorem*{THM*}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem*{lemma*}{Lemma}
\newtheorem*{claim}{Claim}
\newtheorem*{cor}{Corollary}
\newtheorem{prop*}{Proposition}
\newtheorem*{conj}{Conjecture}
\theoremstyle{remark}
\newtheorem*{example*}{Example}
\theoremstyle{definition}
\newtheorem*{definition*}{Definition}
% Special theorem displays
\newenvironment{thm}%
  {\begin{leftbar}\begin{THM}
}{%
  \end{THM}\end{leftbar}
}
\newenvironment{thm*}%
  {\begin{leftbar}\begin{THM*}
}{%
  \end{THM*}\end{leftbar}
}
\newenvironment{prop}%
  {\begin{leftbar}\begin{prop*}
}{%
  \end{prop*}\end{leftbar}
}
\newenvironment{definition}%
  {\begin{leftbar}\begin{definition*}
}{%
  \end{definition*}\end{leftbar}
}
\newenvironment{example}%
  {\begin{leftbar}\begin{example*}
}{%
  \end{example*}\end{leftbar}
}
% Paired math operators
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
% Math helper functions
\newcommand{\congr}[2]{{[#1]}_{#2}}
\newcommand{\set}[1]{\{#1\}}
\newcommand{\genset}[1]{{\langle{}#1\rangle{}}}
\newcommand{\abs}[1]{\lvert{}#1\rvert{}}
\newcommand{\norm}[1]{\lVert{}#1\rVert{}}
\newcommand{\bmx}[1]{\begin{bmatrix}#1\end{bmatrix}}
\newcommand{\pmx}[1]{\begin{pmatrix}#1\end{pmatrix}}
  \newcommand{\sbmx}[1]{{\left[\begin{smallmatrix}#1\end{smallmatrix}\right]}}
\newcommand{\smallxymatrix}[2]{\xymatrix@C-=#1in@R-=#1in{#2}}
\renewcommand\labelitemi{\small$\bullet$}
%
%\makeatletter
\tikzset{%
  gdot/.style={draw,shape=circle,fill=black,inner sep=1},
  ddot/.style={draw,shape=circle,fill=white,inner sep=1},
  dedge/.style={draw=black!40},
  midarrow/.style={postaction={decorate},decoration={markings,mark=at position 0.5 with {\arrow{>}}}}
}
%\makeatother
\begin{document}
%
%
%

\title{Flow polytopes}
\author{Nils Olsson}
\maketitle

Based upon the book \emph{Combinatorial Reciprocity Theorems} by
Matthias Beck and Raman Sanyal; 2010 \emph{Mathematics Subject Classification}.

\section*{Topic:}

@@

\section*{Introduction}

Out of all the various topics in mathematics, graph theory is a particularly
interesting one, and is relevant to a wide variety of other mathematical
topics as well as computer science.
What this paper intends to do is, in the first section, is introduce the topic
of graphs. Many different families of graphs exist, but for this paper we will
concern ourselves with directed graphs, and $\Z_n$-flow functions which act
upon these graphs.

Then, in the second section, we will briefly explore the material of
chapter 7 of Beck and Sanyal's book \emph{Combinatorial Reciprocity Theorems},
and specifically a single main reciprocity expression theorem:
\begin{thm*}[Flow counting function]
  Let $G=(V,E)$ be a graph with a fixed base orientation.
  Then $\varphi_G(n)$ counts the number of nowhere-zero $\Z_n$-flows of $G$,
  and is defined:
  \[
    \varphi_G(n)
    =\sum_{\vec b\in\calC(G)}\sgehr_{\sfP_G^\circ(\vec b)}(n)
    ={(-1)}^{\xi(G)}\sum_{\vec b\in\calC(G)}\sgehr_{\sfP_G(\vec b)}(-n).
  \]
\end{thm*}
What this theorem provides us with is a method of traveling from the topics of
graphs and combinatorics and into the study of a seemingly unrelated
mathematical topic: polytopes.
Specifically, using polytopes to count the number of possible flows
on directed graphs, subject to a few constraints,
and even more specifically we will be looking at a way to count only the
nowhere-zero $\Z_n$-flows of a directed graph.
This should of course require a brief introduction in polytopes as well,
but as time permitted this paper wont include an introducting into polytopes or
Ehrhart reciprocity.

What this paper does \emph{not} intend to do (at least, not anymore) is
thoroughly explain the proofs of this theorem, in all of its intricacies.
Instead the second section is a restructuring of parts of chapter 7 in
Beck and Sanyal's book, in an attempt to provided a clearly progression of
concepts.

Additionally, this paper will \emph{not} work on a specific example of a graph,
as I had originally intended. I had hoped to use a very simple directed
graph to demonstrate some of the underlying mechanics of the propositions and
theorems of the book, but found it to be beyond my own scope.
%Instead, we will use a very specific and very simple example of a graph: one
%with three vertices and an edge between each pair of vertices.
%Ostensibly this forms a triangle, the simplest type of graph to which this
%theorem applies to. We will also chose a specific base orientation.
%It is with this graph that we will develop some intuition into how the main
%theorem works.
\begin{figure}[h]
  \begin{tikzpicture}
    \path[every node/.style={gdot}]
      (-1,0) node (v_1){} --
      ( 1,0) node (v_2){} --
      ($(0,{sqrt(3)})$) node (v_3){} --
      cycle;
    \node[left] at (v_1){$v_1$};
    \node[right] at (v_2){$v_2$};
    \node[above] at (v_3){$v_3$};
    \draw[midarrow] (v_1) -- node (e_1) [below]{$e_1$} (v_2);
    \draw[midarrow] (v_2) -- node (e_2) [above right]{$e_2$} (v_3);
    \draw[midarrow] (v_3) -- node (e_3) [above left]{$e_3$} (v_1);
  \end{tikzpicture}
  \caption{The intended example graph}
\end{figure}

@@

\section{Graphs and flows}

@@

We set off by defining and briefly explaining the topics of both graphs and
flows.
A vertex is most intuitively understood as a coordinate in space.
For instance the coordinates $x=1$ and $y=1$ can define a vertex in
Euclidean, $\R^2$, space.
An edges then is a connection between two vertices.
Edges do not necessarily need to have any particular geometric representation,
but the concept that two edges can cross over one-another will be required
later.
\begin{definition}[Graph]
  A \emph{graph} $G=(V,E)$ is a collection of $V$, a set of vertices, and $E$, a
  set of edges between the vertices.
\end{definition}

Now, a graph does not need to have any vertices.
In fact a graph with no vertices is still a graph, but it is \emph{the} empty
graph, and will not be useful in the development of the main theorem.
Nor does a graph need to have any edges (except that if a graph has edges,
then there must be vertices to which those edges connect).
Also, since a graph is just a collection of vertices and edges, we can have
graphs within graphs, i.e.\ a \emph{subgraph}, which consist of a subset of
the vertices and edges of the whole graph (in particular, every graph
necessarily has the empty graph as a subgraph, and every graph has itself as
a subgraph).

As mentioned before,
vertices, and by extension a graph, can exist within some arbitrary
d-dimensional space. However, for this report we will only concern ourselves
with graphs in two-dimensional, Euclidean space (or at least something
analogous to two dimensional).
(I will also point out that the words ``vertex'' and ``node'' are equivalent.
This paper at times switches between either word, interchangeably.)

So, given a graph with vertices and edges between vertices, we define one
classification of a subset with at least a few of its vertices, and with
possibly a few of its edges:
\begin{definition}[Component, connected component]
  A \emph{connected component} (or simply a \emph{component}) of a graph $G$ is
  a subgraph in which any two
  vertices are either directly connected via an edge, or are connected to
  one-another via a \emph{path} (a series of vertices and edges). A single
  vertex with no connecting edges is in itself a connected component.
\end{definition}

Any graph then is the disjoint union of connected components.
For example Figure~\ref{fig:disjoint-union} can be called a single graph that is the disjoint
union of three distinct connected components.
\begin{figure}[ht]
  \begin{tikzpicture}[every node/.style=gdot]
    \path
      (.5,.5) node (a) {}
      (0,0) node (b) {} (1,0) node (c) {} (1,1) node (d) {} (0,1) node (e) {}
      (2,1) node (f) {} (2,0) node (g) {} (3,0) node (h) {};
    \draw (a);
    \draw (b) -- (c) -- (d) -- (e) -- (b);
    \draw (f) -- (g) -- (h);
  \end{tikzpicture}
  \caption[Disjoin union]{}\label{fig:disjoint-union}
\end{figure}
Traversing from one vertex of a graph to another by way of edges only, we see
that it may be impossible to travel to all of its vertices, hence a graph
consisting of potentially disjoint components.
\begin{lemma*}
  A connected component of the graph $G$ is a maximal subgraph of $G$ in which
  any two nodes are connect by a path.
\end{lemma*}
\begin{lemma*}
  A graph $G$ is \emph{connected} if it has only one connected component.
\end{lemma*}
\begin{definition}[Bridge]
  A \emph{bridge} is a single edge which serves as a connection between two connected
  components which, if removed, would increase the number of connected
  components of the graph by one.
\end{definition}

For example, removing the bridge in Figure~\ref{fig:bridge} turns a single
connected component into two disjoint connected components:
one complete graph into a graph with two complete subgraphs.
\begin{figure}[ht]
  \begin{tikzpicture}[every node/.style=gdot]
    \path
      (-.5,.5) node (a) {} ($({-(1+sqrt(3))/2},0)$) node (b) {} ($({-(1+sqrt(3))/2},1)$) node (c) {}
      (.5,.5) node (d) {} ($({ (1+sqrt(3))/2},0)$) node (e) {} ($({ (1+sqrt(3))/2},1)$) node (f) {};
    \draw (a) -- (b) -- (c) -- (a);
    \draw (d) -- (e) -- (f) -- (d);
    \draw[dashed] (a) -- (d);
  \end{tikzpicture}
  \hspace{1in}
  \begin{tikzpicture}[every node/.style=gdot]
    \path
      (-.5,.5) node (a) {} ($({-(1+sqrt(3))/2},0)$) node (b) {} ($({-(1+sqrt(3))/2},1)$) node (c) {}
      (.5,.5) node (d) {} ($({ (1+sqrt(3))/2},0)$) node (e) {} ($({ (1+sqrt(3))/2},1)$) node (f) {};
    \draw (a) -- (b) -- (c) -- (a);
    \draw (d) -- (e) -- (f) -- (d);
  \end{tikzpicture}
  \caption[Bridges]{An undirected graph
  with and without a bridge.}\label{fig:bridge}
\end{figure}

@@

So far we've only looked at \emph{unoriented} graphs, but
now we introduce a notion of direction, or orientation, to graphs.
\begin{definition}[Directed, oriented graph]
  A \emph{directed} or \emph{oriented graph} is a graph on which we
  define an \emph{orientation} via a subset $\rho$ of the edges $E$ of the
  graph. For each edge $e=v_i v_j\in E$ (connecting the vertices $v_i,v_j\in
  V$) with $i<j$ we direct:
  \[
    v_i\stackrel{e}{\leftarrow}v_j\text{ if }e\in\rho\text{, and }
    v_i\stackrel{e}{\rightarrow}v_j\text{ if }e\notin\rho.
  \]
  The graph $G$ with orientation $\rho$ is denoted ${}_\rho G$.
\end{definition}
This definition may be confusing at first glance because $\rho$
may be missing some of the edges of $E$. It only includes the edges with
their direction pointing from the vertex of higher index to the vertex of lower
index (which does certainly feel backwards).
Then, all of the edges not in $\rho$ are directed in the opposite
direction: from lower to higher index.

Figure~\ref{fig:directed-graph} demonstrate how this encapsulation works by
applying an orientation $\rho=\{14,23,24\}$ onto a particular graph in three
steps.
\begin{figure}[ht]
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style=gdot]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw (v_4) -- (v_3) -- (v_2) -- (v_1) -- (v_4) -- (v_2);
    \end{tikzpicture}
    \caption{The undirected graph.}
  \end{subfigure}
  %\hspace{1in}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style=gdot]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw[midarrow] (v_4) -- (v_1);
      \draw[midarrow] (v_4) -- (v_2);
      \draw[midarrow] (v_3) -- (v_2);
      \draw (v_3) -- (v_4);
      \draw (v_1) -- (v_2);
    \end{tikzpicture}
    \caption{$\rho$ edges marked.}
  \end{subfigure}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style={gdot}]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw[midarrow] (v_4) -- (v_1);
      \draw[midarrow] (v_4) -- (v_2);
      \draw[midarrow] (v_3) -- (v_2);
      \draw[midarrow] (v_3) -- (v_4);
      \draw[midarrow] (v_1) -- (v_2);
    \end{tikzpicture}
    \caption{$E\setminus\rho$ edges marked.}
  \end{subfigure}
  \caption[A directed graph]{A directed graph.}\label{fig:directed-graph}
\end{figure}

As an aside, there are many algorithms in computer science pertaining to graphs,
some for finding optimal ways to traverse the paths of graphs. In some
models each edge is considered to be of equal ``weight'' to any
other. That is, if two unweighted edges both originate from a node $v_1$
and both point to a second node $v_2$, then there is no reason to choose one
edge over the other in traveling from $v_1$ to $v_2$.
On the other hand, we might assign a weight to
each edge, presenting a cost-benefit analysis in choosing a path to take (as in
Figure~\ref{fig:path-finding}, where the edge with weight 5 now becomes most
``expensive'' to traverse versus the one with weight 2).
\begin{figure}[ht]
  \begin{tikzpicture}
    \path (0,0) node[gdot] (a){} (2,0) node[gdot] (b){};
    \draw[midarrow] (a) to [out=45, in=135] node[above]{} (b);
    \draw[midarrow] (a) to [out=315,in=225] node[below]{$\phantom 5$} (b);
  \end{tikzpicture}
  \hspace{1in}
  \begin{tikzpicture}
    \path (0,0) node[gdot] (a){} (2,0) node[gdot] (b){};
    \draw[midarrow] (a) to [out=45, in=135] node[above]{$2$} (b);
    \draw[midarrow] (a) to [out=315,in=225] node[below]{$5$} (b);
  \end{tikzpicture}
  \caption{A directed graph, with and without weights.}\label{fig:path-finding}
\end{figure}

\begin{definition}[Directed path, directed cycle]
  A \emph{directed path} in ${}_\rho G$ is a sequence $v_0,v_1,\ldots,v_s$ of
  distinct nodes such that $v_{j-1}\to v_j$ is a directed edge in ${}_\rho G$
  for all $j=1,\ldots,s$. If $v_s\to v_0$ is also a directed edge, then
  $v_0,v_1,\ldots,v_s,v_{s+1}=v_0$ is called a \emph{directed cycle}.
\end{definition}

\begin{definition}[Acyclic orientations]
  An orientation $\rho$ of $G$ is acyclic if there are no directed cycles in
  ${}_\rho G$.
\end{definition}

Dual to the notion of acyclic orientations are orientations which are
\emph{totally cyclic}, where every edge in ${}_\rho G$ is contained in a
directed cycle.
\begin{figure}[h]
  \begin{tikzpicture}
    \path[every node/.style={gdot}]
      (-1, 0) node(a){} --
      (-1, 1) node(b){} --
      ( 0, 1) node(c){} --
      ( 0, 0) node(d){} --
      ( 1, 0) node(e){} --
      ( 1, 1) node(f){};
    \draw[midarrow] (c) -- (d);
    \draw[midarrow] (a) -- (b);
    \draw[midarrow] (b) -- (c);
    \draw[midarrow] (d) -- (a);
    \draw[midarrow] (e) -- (f);
    \draw[midarrow] (f) -- (c);
    \draw[midarrow] (d) -- (e);
  \end{tikzpicture}
  \caption{A totally cyclic graph with two cycles.}
\end{figure}

\begin{definition}[Cyclotomic number]
  We define the \emph{cyclotomic number} of $G$ as:
  \[
    \xi(G)=|E|-|V|+c.
  \]
  Where $c=c(G)$ is the number of connected components of $G$.
\end{definition}

Another topic of graphs is coloring, where every vertex is assigned a color
value, and a proper coloring is where no vertices connected by an edge share the
same color. Directions and cycles provide a particular way to induce a coloring
onto the vertices of a graph, but this is outside the scope of this paper.
However, acyclic and cyclic orientations are a critical component of
the main focus of this paper, the nowhere-zero $\Z_n$-flow counting function,
which we now introduce.

@@

  \caption{Various graphs and their duals.}\label{fig:graphs-and-duals}

@@

% ------------------------------------------------------------------------------
% 7.6
% ------------------------------------------------------------------------------
\hrulefill

\section{Reciprocity in counting nowhere-zero $\Z_n$-flows}

Originally my intention of this second section of the paper was to demonstrate,
by example, the underlying mechanisms of the theorems and propositions in the
Beck and Sanyal book. Instead what I found was that my grasp on the topic was
shaky at best. What proved the most useful, to me at least, was to piece
together a different order of the book's chapter 7 definitions, propositions and
theorems; an order which feels much more natural in moving from a graph and
combinatorial centric topic of flows, into a more polytope centric topic of
reciprocity expressions.
It was also at this time that I recognized that many topic introduced in the
first second of this paper would now have no connection to this second section,
specifically dual graphs and cyclic/acyclic orientations.
They are in fact critical in the proofs of the propositions behind the
$\Z_n$-flow counting function, but by limitations of time wont find any use
here-on-out.
